### 2023-05-08 Documentation/UFMG/PFCII/Results/Results.tex
documentation

In this chapter, we will present the results obtained from our project. 
Firstly, we will demonstrate how the automation routine enhanced the research experiments. 
Next, we will discuss the validation of the classification results. 
Finally, we will showcase the performance of the implemented controller algorithms.

@@

The automated experiment routine is capable of acquiring a significantly more precise and extensive amount of data than what can be achieved by a human. 
With the save data by streaming in saving thread \ref{subsec:save_thread}, we achieved less program memory during the experiment and, specially, safety for not losing the data in case the program fail during an experiment.

Also, most power supplies rely on manual potentiometer adjustment to select the set point, which results in imprecise potential selection by a person and takes some time to achieve the desired voltage.
Moreover, the electrospray phenomena has a known hysteresis, that can be seen in figure \ref{fig:ganan_calvo_fig}, and can perform different results depending on the previous electric potential.
By running a computer routine that automatically sends commands to the power supply and pump machine we can achieve faster and more reliable experiment data points.

@@

\begin{figure}[H]
    \center
    \includegraphics[width=16cm]{Figuras/19:03/axs1.png}
    \caption{Print screen of the window shows user interface during the experiment.
        We can see the image generated by the camera in the background.
        The routine code running in pycharm software on the right side.
        And also real time signal plottings of the current data on the left side.}
        \label{fig:multi_class_exp1}
\end{figure}

@@

The classification results in our algorithm implemented as described in section \ref{sec:section_classification} showed good accuracy for all the classifications described in section \ref{sec:spraying_modes_subsec}.
Nevertheless, the multi jet classification made by a current value factor of 1.14 above the cone jet is just effective for pure ethanol, liquid used in all this project.

@@

We will categorize our classification results into two main groups - the step routine and the map routine. These two routines have showed valuable insights and comprehensible results of the electrospraying process.

@@

Our first classification results were made exploring the voltage ranges. 
For that, we fixed a flow rate to 0.7 uL/min. and ran a step routine as defined in \ref{subsec:step_routine}. 
The figure \ref{fig:step_class} shows three graphs.

The first shows the controller output signal as an input voltage of the process. As it is a step routine we implemented a increasing voltage with steps of sizes 50V and time between each step of 5 seconds. The voltage range is between 3K~10K Volts.

The second is the raw output data collected by the oscilloscope in \emph{data\_acquisition\_thread()}. 
The sampling rate is 10KHz. Therefore, this experiment of 700s has 7 Million data points just of current data. 
This is an example of how scalable the data collected can be depending on the experiment time. This will be even more noticeable in mapping experiments.

@@

For example, dripping mode has a current mean of 0V. The Intermittent state has a high variation of values that can be seen by the increased thickness on the graph. The Cone jet is a thinner graph because of its constant signal. Multi Jet  has the same shape as Cone jet but with a higher mean value. Corona sparks are not showed in the graph because its discharges has a high current value above the axis limits.

@@

For validation with literature and also to expose the benefits of the automated routine and classification, the map sequence proof itself the best result of this work.

    Initially, for better understand pure ethanol classification regions through voltage and flow rate ranges, I made a manual map seen in figure \ref{fig:stability_1}.

@@

    In order to validate our automatic classification, experiments were made comparing both visual and automatic stability island on the same experiment.
    Figure \ref{fig:stability_2} shows that automatic stable cone jet region could be identified in the same region as visually seen by the high speed camera (data acquired manually).

@@

    With the development of a Multi Jet classification using the logic explained in section \ref{sec:section_classification}, I repeated the same experiment as shown in Figure \ref{fig:stability_6}.

@@

    To have a better comparison with literature, specifically the Gañán-Calvo\cite{gananCalvo} stability islands showed in figure \ref{fig:ganan_calvo_fig}, with a visual juxtaposition of the shapes, we displayed in figure \ref{fig:stability_8} the data using the non-dimensional numbers used in his work\cite{gananCalvo}.

@@

    The results of mapping experiment corresponded visual, literature and automatic regions of classification, giving authentication to our method.

@@

    The results of the classification process were favorable for experiment automation, however, the inaccuracy of the classification by statistical methods, specially the Multi Jet\ref{subsec:Multi Jet}, limited the development of a complete control project.
    Together with the amount of variables that need to be syntonized makes it hard to stabilize in a desired mode.

    Even with all those problems we could implement a simple controller project, described in \ref{sec:control_model}, that validated the efforts of remodeling all software into a closed loop control model.

    The controller not only stabilized in Cone Jet mode, but also could reject perturbation, as seen in Figure \ref{fig:control_results}, where I manually changed the flow rate.
    First graph shows current acquired. It starts as intermittent state, with an oscillating signal. We can validate this with the blue colored sample on the third graph representing its automatic classification. 
    The second graph shows the controller actuating increasing the voltage to achieve a stable cone jet mode.
    After a period of time in cone jet, I manually increased the flow rate, serving as a perturbation of the system, and we can see that the effect of that was to return to intermittent state with the oscillating signal. The controller again, actuate increasing the voltage to reach the cone jet stability island. 
    This procedure was repeated after some seconds of stabilization in cone jet, increasing again the flow rate manually, and we can see the system automatically adjusting the voltage.
        \begin{figure}[H]
            \center
            \includegraphics[width=16cm]{Figuras/19:03/control_first_results.png}
            \caption{Simple controller validation experiment. For this experiment we waited the system to stabilize in the desired spraying mode. After that we made a perturbation test changing the flow rate. The classification legend follows the same color reference in all project.}
            \label{fig:control_results}
        \end{figure}

    After 200s of experiment we had some breaks up of the cone jet into a droplet and stabilized again. This is represented by the peak values in current signal. The classification correctly identified as intermittent, seen as blue dots in the third graph, and the controller reacted increasing the voltage, in a situation that can be considered as noise. A filter can be applied in a more complex controller project to avoid this.

@@

In this chapter we exposed the results of this project. The automatic routine, real time classification and control were all achieved and implemented. However, they can all be improved. In the next Chapter I will conclude this document with discussions about the results and proposal of continuation.

>>>>

This chapter will present the results obtained from this project. 
Firstly, it will be demonstrated how the automation routine enhanced the research experiments. 
Next, a discussion on the validation of the classification results. 
Finally, a showcase of performance on the implemented controller algorithms.

@@

The automated experiment routine is capable of acquiring a significantly more precise and extensive amount of data than what can be achieved by a human. 
With the save data by streaming in saving thread \ref{subsec:save_thread}, it was achieved less program memory during the experiment and, specially, safety for not losing the data in case the program fail during an experiment.

Also, most power supplies rely on manual potentiometer adjustment to select the set point, which results in imprecise potential selection by a person and takes some time to achieve the desired voltage.
Moreover, the electrospray phenomena has a known hysteresis, that can be seen in figure \ref{fig:ganan_calvo_fig}, and can perform different results depending on the previous electric potential.
Running a computer routine that automatically sends commands to the power supply and pump machine allowed obtaining faster and more reliable experiment data points.

@@

\begin{figure}[H]
    \center
    \includegraphics[width=16cm]{Figuras/19:03/axs1.png}
    \caption{Print screen of the window shows user interface during the experiment.
        The image generated by the high speed camera can be seen in the background.
        The routine code running in pycharm software on the right side.
        And also real time signal plotting of the current data on the left side.}
        \label{fig:multi_class_exp1}
\end{figure}

@@

The classification results in the algorithm implemented as described in section \ref{sec:section_classification} showed good accuracy for all the classifications described in section \ref{sec:spraying_modes_subsec}.
Nevertheless, the multi jet classification made by a current value factor of 1.14 above the cone jet is just effective for pure ethanol, liquid used in all this project.

@@

The classification result will be categorized into two main groups - the step routine and the map routine. These two routines have showed valuable insights and comprehensible results of the electrospraying process.

@@

The first classification results were made exploring the voltage ranges. 
For that, a flow rate was fixed to 0.7 uL/min. and ran a step routine as defined in \ref{subsec:step_routine}. 
The figure \ref{fig:step_class} shows three graphs.

The first shows the controller output signal as an input voltage of the process. As it is a step routine, an increasing voltage with steps of sizes 50V and time between each step of 5 seconds was implemented. The voltage range is between 3k-10k Volts.

The second is the raw output data collected by the oscilloscope in \emph{data\_acquisition\_thread()}. 
The sampling rate is 100kHz. Therefore, this experiment of 700s has 70 Million data points just of current data. 
This is an example of how scalable the data collected can be depending on the experiment time. This will be even more noticeable in mapping experiments.

@@

For example, dripping mode has a current mean of 0V. The Intermittent state has a high variation of values that can be seen by the increased thickness on the graph. The cone jet is a thinner graph because of its constant signal. Multi Jet  has the same shape as cone jet but with a higher mean value. Corona sparks are not showed in the graph because its discharges has a high current value above the axis limits.

@@

For validation with literature and also to expose the benefits of the automated routine and classification, the map sequence proof itself the best result of this work. 
Initially, for better understand pure ethanol classification regions through voltage and flow rate ranges, it was made a manual map seen in figure \ref{fig:stability_1}.

@@

    In order to validate the automatic classification, experiments were made comparing both visual and automatic stability island on the same experiment.
    Figure \ref{fig:stability_2} shows that automatic stable cone jet region could be identified in the same region as visually seen by the high speed camera (data acquired manually).

@@

    With the development of a multi jet classification using the logic explained in section \ref{sec:section_classification}, the same experiment was repeated, as shown in Figure \ref{fig:stability_6}.

@@

    To have a better comparison with literature, specifically the Gañán-Calvo\cite{gananCalvo} stability islands showed in figure \ref{fig:ganan_calvo_fig}, with a visual juxtaposition of the shapes, the data displayed in Figure \ref{fig:stability_8} using the non-dimensional numbers used in his work\cite{gananCalvo}.

@@

    The results of mapping experiment corresponded visual, literature and automatic regions of classification, giving authentication to the method.

@@

    The results of the classification process were favorable for experiment automation, however, the inaccuracy of the classification by statistical methods, specially the Multi Jet\ref{subsec:Multi Jet}, limited the development of a complete control project.
    Together with the amount of variables that need to be syntonized makes it hard to stabilize in a desired mode.
    Even with all those problems, a simple controller project could be implemented, as described in section \ref{sec:control_model}, that validated the efforts of remodeling all software into a closed loop control model.

    The controller not only stabilized in cone jet mode, but also could reject perturbation, as seen in Figure \ref{fig:control_results}, where the flow rate was manually changed.
    First graph shows current acquired. It starts as intermittent state, with an oscillating signal. This can be validated with the blue colored sample on the third graph representing its automatic classification. 
    The second graph shows the controller actuating increasing the voltage to achieve a stable cone jet mode.
    After a period of time in cone jet, a manual increase in the flow rate was done, serving as a perturbation of the system, and it can be seen that the effect of that was to return to intermittent state with the oscillating signal. The controller again, actuate increasing the voltage to reach the cone jet stability island. 
    This procedure was repeated after some seconds of stabilization in cone jet, increasing once more the flow rate manually, and again the system automatically adjusting the voltage.
        \begin{figure}[H]
            \center
            \includegraphics[width=16cm]{Figuras/19:03/control_first_results.png}
            \caption{Simple controller validation experiment. For this experiment, after waiting the system to stabilize in the desired spraying mode it was made a perturbation test changing the flow rate. The classification legend follows the same color reference in all project.}
            \label{fig:control_results}
        \end{figure}

    After 200s of experiment the system had some breaks up of the cone jet into a droplet and stabilized again. This is represented by the peak values in current signal. The classification correctly identified as intermittent, seen as blue dots in the third graph, and the controller reacted increasing the voltage, in a situation that can be considered as noise. A filter can be applied in a more complex controller project to avoid this.

@@

In this chapter the results of this project were exposed. The automatic routine, real time classification and control were all achieved and implemented. However, they can all be improved. Next Chapter will conclude this document with discussions about the results and proposal of continuation.

### 2023-05-08 Documentation/UFMG/PFCII/systemDescription/systemDescription.tex
documentation

In this chapter, we will discuss the equipment used in detail, including its integration and how it was modeled and implemented in both hardware and software. 
The python automation routine that is the focus of this project was started by a previous student\cite{Monica}, and the software itself was originally created as an electrospray multipurpose library.

@@

    The software has an interface to integrate the HVPS to our routine. This interface can be found in \emph{FUG\_function.py} file where is located the functions used to control and collect data from this instrument.
    In case of change of equipment, a new interface must be created within this file to match another manufacturer specifications.

@@

    Wireless communication allows us to make measurements disconnected to an external power supply, which increase safety when using high voltage potential references and also reduce the signal noise collected from external power lines.
    The current is routed directly via the input, hence the oscilloscope measures the voltage dropped via its input resistance (which can be switched between 1 or 2 Mohms).
    TiePie WifiScope WS6 has a resolution of up to 16 bit at a minimal input range of 200 mV, sufficient to measure currents down to 1 nA.

@@

  The pump integration in the automation algorithm was done in this work, bringing us a new controllable variable, the flow rate. We can control the spraying mode with the
  two main variables that affect the system. 
  It brings more complexity for the system since now we are dealing with multivariable control.
  Controlling also the flow rate gives to this project a new dimension in the system giving us freedom to explore the flow rate properties.
  With this new input variable our control is a MISO (Multiple Inputs Single Output) system.

  About the pump interface, as there was no good ready-to-use library, I developed a simple and intuitive interface to be used in our software routine.
  The communication protocol used is RS-232 and the pump commands list were found in the user manual.

@@

In figure \ref{fig:setup} we illustrate a diagram with all the key components of our system. The peripherals are connected to a computer running the software routine via serial communication. 
This diagram encapsulates our process system and will be used as a sub-system in our control model. The input of our process system are power supply voltage and pump machine flow rate, referred here as the controller values. The output is the oscilloscope current sample.

@@

There are also many minor variables that affects the experiment stability. More details are explained in \ref{sec:setup_validation}.

@@

The software was reformulated on top of the control model represented in figure \ref{fig:control_model_fig}. The process of the control loop is the same as represented in \ref{fig:setup}. 
Each other subsystem in the control loop is a separate thread that will be explained bellow.

@@

    In order to implement this closed loop control model to the software and explore parallel processing, each sub-system was implemented as a separate Thread.
    For concurrency on flux of data between threads was used queues structures.
    A queue is an abstract data type that holds an ordered, linear sequence of items. You can describe it as a first in, first out (FIFO) structure.

@@

        It is responsible for sending the power supply set voltage values and the syringe pump the flow rate set values according to the sequence selected.
        Also, responsible for sending the finish event command that end the routine and trigger the threads to close their routines.
        As input, we have the setup config file and the \emph{feedback\_queue}. As output, we have the values in the \emph{controller\_output\_queue()}.

@@

        It is responsible for reading the current data from the oscilloscope, humidity and temperature data from the DHT11 sensor, voltage from the power supply, flow rate from the pump and concatenate into one sample data.
        As output, we have the values in the \emph{data\_queue()}.

@@

        It is responsible for calculating the statistical values from the raw data and classify it in the respective spray mode for that sample.
        As output, we have the values in the \emph{save\_data\_queue()}, \emph{plotting\_queue()} and \emph{feedback\_queue()}.

@@

        After processing, the data is saved in real time in a json file using \emph{jsonstreams} library to save one sample structure at a time.
        With the new streaming model of saving a new structure of the collected data were created.
        Instead of having all data measurements values and after all data processing values we now are saving for each sample the measurements and processing values.
        The data acquired in each sample of 0.5s is shown in figure \ref{fig:data_sample}.

@@

        This new way of saving the measurement and processing data altogether separate by enumerated samples prevents loss of data in case of error during experiments and, specially improve in the usability of the data for further deep analysis. One experiment of around 1 hour collects around 6 GB of data. Organizing that amount of data to make it fast and easy to use was an important improvement in the software.
        To work and analyze the data we can use  pandas Data frame. A python widely used framework for data analysis.
        With the command:

@@

        The only running function that should run in the main thread because the plotting library \emph{matplotlib} incompatibilities of running outside the main function. 
        It is responsible for plotting in real time the current sample acquired, and its respective fast Fourier transform to evaluate the sample frequency spectrum.
        It takes as input the \emph{plot\_data\_queue} from the \emph{processing\_thread()} and displays three graphs updated in real time on the screen during the experiment.

@@

    In this chapter we described the system description detailing the hardware instruments, how it was modelled and its implementation in the software.

>>>>

In this chapter, will be discussed the equipment used in detail, including its integration and how it was modeled and implemented in both hardware and software. 
The python automation routine that is the focus of this project was started by a previous student\cite{Monica}, and the software itself was originally created as an electrospray multipurpose library.

@@

    The software has an interface to integrate the HVPS to the routine. This interface can be found in \emph{FUG\_function.py} file where is located the functions used to control and collect data from this instrument.
    In case of change of equipment, a new interface must be created within this file to match another manufacturer specifications.

@@

    Wireless communication allows measurements disconnected to an external power supply, which increase safety when using high voltage potential references and also reduce the signal noise collected from external power lines.
    The current is routed directly via the input, hence the oscilloscope measures the voltage dropped via its input resistance (which can be switched between 1 or 2 Mohms).
    TiePie WifiScope WS6 has a resolution of up to 16 bit at a minimal input range of 200 mV, sufficient to measure currents down to 1 nA.

@@

  The pump integration in the automation algorithm was done in this work, bringing a new controllable variable, the flow rate. The spraying mode can be controlled using the two main variables that affect the system. 
  It brings more complexity for the system since now it is a multivariable control.
  Controlling also the flow rate provide this project a new dimension in the system giving freedom to explore the flow rate properties.
  With this new input variable the control system is MISO (Multiple Inputs Single Output).

  About the pump interface, as there was no good ready-to-use library, it was developed a simple and intuitive interface to be used in the software routine.
  The communication protocol used is RS-232 and the pump commands list were found in the user manual.

@@

Figure \ref{fig:setup} illustrate a diagram with all the key components in the system. The peripherals are connected to a computer running the software routine via serial communication. 
This diagram encapsulates the process system and will be used as a sub-system of the control model. The input in the process are power supply voltage and pump machine flow rate, referred here as the controller values. The output is the oscilloscope current sample.

@@

There are also many minor variables that affects the experiment stability. More details are explained in appendix \ref{sec:setup_validation}.

@@

The software was reformulated on top of the control model represented in Figure \ref{fig:control_model_fig}. The process of the control loop is the same as represented in Figure \ref{fig:setup}. 
Each other subsystem in the control loop is a separate thread that will be explained bellow.

@@

    In order to implement this closed loop control model to the software and explore parallel processing, each sub-system was implemented as a separate thread.
    For concurrency on flux of data between threads was used queues structures.
    A queue is an abstract data type that holds an ordered, linear sequence of items. You can describe it as a first in, first out (FIFO) structure.

@@

        The controller thread is responsible for sending the power supply set voltage values and the syringe pump the flow rate set values according to the sequence selected.
        Also, responsible for sending the finish event command that end the routine and trigger the threads to close their routines.
        As input, there is the setup config file and the \emph{feedback\_queue}. As output, the values in the \emph{controller\_output\_queue()}.

@@

        The data acquisition thread is responsible for reading the current data from the oscilloscope, humidity and temperature data from the DHT11 sensor, voltage from the power supply, flow rate from the pump and concatenate into one sample data.
        As output, the values in the \emph{data\_queue()}.

@@

        The data processing is responsible for calculating the statistical values from the raw data and classify it in the respective spray mode for that sample.
        As output, the values in the \emph{save\_data\_queue()}, \emph{plotting\_queue()} and \emph{feedback\_queue()}.

@@

        After processing, the data is saved in real time in a json file using \emph{jsonstreams} library to save one sample structure at a time.
        With the new streaming model of saving, instead of having all data together in program memory to be saved in the end, it is saving each sample in real time during the experiment.
        The data acquired in each sample of 0.5s is shown in Figure \ref{fig:data_sample}.

@@

        This new way of saving the measurement and processing data altogether separate by enumerated samples prevents loss of data in case of error during experiments and, specially improve in the usability of the data for further deep analysis. One experiment of around 1 hour collects around 6 GB of data. Organizing that amount of data to make it fast and easy to use was an important improvement in the software.
        To work and analyze the data it can be used pandas Data frame. A python widely used framework for data analysis.
        With the command:

@@

        The only running function that should run in the main thread because the plotting library \emph{matplotlib} incompatibilities of running outside the main function. 
        The plotting function is responsible for plotting in real time the current sample acquired, and its respective fast Fourier transform to evaluate the sample frequency spectrum.
        It takes as input the \emph{plot\_data\_queue} from the \emph{processing\_thread()} and displays three graphs updated in real time on the screen during the experiment.

@@

    In this chapter it is described the system description detailing the hardware instruments, how it was modelled and its implementation in the software.
